Computing astrophysical inferences with diffusive nested sampling

### 2014-02-23 ms.tex
A figure

\subsection{MCMC inside nested sampling}
The main challenge in Nested Sampling is to generate new points from the
constrained prior distribution $\propto \pi(\pars)\mathds{1}(L(\pars) > L^*)$.
Two popular strategies are in use. The first is to use the remaining ``live
points'', which are known to be above the likelihood threshold $L^*$, to build
an approximation to the constrained prior, which can then be used to generate
a new point via rejection sampling \citep{multinest}.

The second common strategy is to use MCMC. Since the remaining live points are
known to exceed the likelihood threshold, they can be used to initialise an
MCMC run. This strategy is very simple but has known problems. For example,
suppose the likelihood function has two separate modes, and it impossible for
MCMC to mix between the two modes. Suppose also that NS is being run with two
points, and there is one point in each mode. The worst point will be discarded
(and its information saved elsewhere), and the new point will be generated by
copying the other point and running MCMC. However, even though the two modes
are completely equivalent, the copying operation will remove one of the modes,
and it will never return.

@@

The zeroth component of this mixture is just the prior $p(\pars)$. The first
mixture component is the prior, but restricted to the domain where the
likelihood is greater than a threshold $L^*_1$. The second mixture component
is the prior, but restricted to the domain where the likelihood is above
$L^*_2$. The likelihood thresholds $\{ L^*_i \}$ are chosen so each
successive constrained prior occupies approximately $\exp(-1)$ times as much
prior mass (``volume'') as the previous constrained prior. This is strongly
related to the algorithm known as ``simulated tempering''
\citep{simulated_tempering}, but uses NS's
sequence of intermediate distributions instead of an annealed sequence.

>>>>

\subsection{MCMC inside nested sampling}
The main challenge in Nested Sampling is to generate new points from the
constrained prior distribution $\propto \pi(\pars)\mathds{1}(L(\pars) > L^*)$.
Two popular strategies are in use. The first is to use the remaining ``live
points'', which are known to be above the likelihood threshold $L^*$, to build
an approximation to the constrained prior, which can then be used to generate
a new point via rejection sampling, as implemented in the
{\tt MultiNest} program \citep{multinest}.

The second common strategy is to use MCMC. Since all of the Nested Sampling
points are
known to exceed the likelihood threshold (which is defined using the lowest
point), they can be used to initialise an
MCMC run. This strategy is very simple to implement but can have
serious problems. For example,
suppose the likelihood function has two separate modes, and it impossible for
MCMC to mix between the two modes. Suppose also that NS is being run with two
points, and there is one point in each mode. The worst point will be discarded
(and its information saved elsewhere), and the new point will be generated by
copying the other point and running MCMC. However, even though the two modes
are completely equivalent, the copying operation will remove one of the modes,
and it will never return.

@@

The zeroth component of this mixture is just the prior $p(\pars)$. The first
mixture component is the prior, but restricted to the domain where the
likelihood is greater than a threshold $L^*_1$. The second mixture component
is the prior, but restricted to the domain where the likelihood is above
$L^*_2$. The likelihood thresholds $\{ L^*_i \}$ are chosen so each
successive constrained prior occupies approximately $\exp(-1)$ times as much
prior mass (``volume'') as the previous constrained prior. This is strongly
related to the algorithm known as ``simulated tempering''
\citep{simulated_tempering}, but uses NS's
sequence of intermediate distributions instead of an annealed sequence.
Retaining ``old'' distributions in the mixture allows
